### 2025-05-08 project3/draft.tex
final draft

\begin{algorithm}
    \caption{Multi-task Learning Framework for \gls{lstm} Metamodels}
    \begin{algorithmic}[1] \label{alg3:multiTaskLearning}
        \STATE \textbf{Input:} learning rate $\alpha$, set of $K$ tasks $\{\mathcal{T}_k\}_{k=1}^K$ with datasets $\mathcal{D}_k = \{(X_k^{(i)}, L_k^{(i)})\}_{i=1}^{M_k}$, task-specific parameters $\theta_k$ for each task $k$, and shared parameters $\theta_0$.

@@

\begin{algorithm}
    \caption{Transfer Learning Framework for \gls{lstm} Metamodels: Combining Fine-tuning, Layer Freezing, and Multi-task Learning}
    \begin{algorithmic}[1] \label{alg3:combined}
        \STATE \textbf{Input:} Set of $K$ tasks $\{\mathcal{T}_k\}_{k=1}^K$, with datasets $\mathcal{D}_k = \{(X_k^{(i)}, L_k^{(i)})\}_{i=1}^{M_k}$ for each task $k$, target dataset $\mathcal{D}_{\text{Ta}}$, learning rate $\alpha_{\text{So}}$ and $\alpha_{\text{Ta}}$, shared parameters $\theta_0$, task-specific parameters $\theta_k$ for each task $k$.

@@

\begin{table}[ht!] 
    \centering
    \begin{tabular}{lcccc} 
    \toprule
    \textbf{Contract} & \textbf{Asset Model} & \textbf{Lapse} & \textbf{$M_{\text{So}}$}  & \textbf{$M_{\text{Ta}}$}\\
    \midrule
    \gls{gmmb} & \gls{rsgbm} & No lapse & $\num{50000}$ & N/A \\
    \gls{gmmb} & \gls{rsgbm} & Static lapse & $\num{50000}$ & $\num{2000}$ \\
    \gls{gmmb} & \gls{rsgbm} & Dynamic lapse & $\num{50000}$ & $\num{2000}$ \\
    \gls{gmwb} & \gls{rsgbm} & Dynamic lapse & N/A & $\num{2000}$ \\
    \bottomrule
    \end{tabular}
    \caption{\gls{va} Contracts for Transfer Learning Experiments}
    \label{tab3:contracts}
\end{table}

@@

All \gls{lstm} metamodels are evaluated based on their training history graphs for the target task, which plot training and validation MSE against the number of training epochs.
The training history graphs visualize the learning curves of the \gls{lstm} metamodels, which provide insights into stability and convergence behavior the metamodels.
We measure the generalization performance of the \gls{lstm} metamodels using the true MSE, which quantifies the accuracy of the \gls{lstm} metamodels in approximating the true inner simulation model of \gls{va} contracts.
The true MSE is computed by comparing the metamodel predictions with the true contract losses, which are approximated by the standard nested simulation procedure with $\num{100000}$ inner replications.
Hyperparameters and network architectures for the \gls{lstm} metamodels are kept the same as the \gls{lstm} metamodel in Section~\ref{sec2:numerical}. 
Due to a limited amount of computational resources, macro replications are not feasible for the \gls{tl} experiments.

@@

\begin{figure}[ht!]
    \centering
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{./project3/figures/figure1a.png}
        \caption{Extensive Training on Target Task} 
        \label{subfig3-1:extensive}
    \end{subfigure}\hfill
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{./project3/figures/figure1b.png}
        \caption{Without \gls{tl}}
        \label{subfig3-1:without}
    \end{subfigure}
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{./project3/figures/figure1c.png}
        \caption{With Fine-tuning}
        \label{subfig3-1:fineTuning}
    \end{subfigure}\hfill
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{./project3/figures/figure1d.png}
        \caption{With Layer Freezing}
        \label{subfig3-1:layerFreezing}
    \end{subfigure}
    \caption{Metamodel performance on \gls{rsgbm} \gls{gmmb} with static lapse}
    \label{fig3:figure1}
\end{figure}

@@

\begin{figure}[ht!]
    \centering
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{./project3/figures/figure2a.png}
        \caption{Fine-tuning from No Lapse} 
        \label{subfig3-2:fromNolapse}
    \end{subfigure}\hfill
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{./project3/figures/figure2b.png}
        \caption{Fine-tuning from Static Lapse}
        \label{subfig3-2:fromLapse}
    \end{subfigure}
    \caption{Fine-tuned Metamodel performance on \gls{rsgbm} \gls{gmmb} with a dynamic lapse}
    \label{fig3:figure2}
\end{figure}

@@

\begin{figure}[ht!]
    \centering
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{./project3/figures/figure3a.png}
        \caption{Freezing LSTM Layers} 
        \label{subfig3-3:freezeLSTM}
    \end{subfigure}\hfill
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{./project3/figures/figure3b.png}
        \caption{Freezing the Fully Connected Layer}
        \label{subfig3-3:freezeFC}
    \end{subfigure}
    \caption{Layer Freezing on \gls{rsgbm} \gls{gmmb} with dynamic lapse}
    \label{fig3:figure3}
\end{figure}

@@

\begin{table}[ht!]
    \centering
    \begin{tabular}{lllll}
    \toprule
    \textbf{Lapse Type} & \textbf{Extensive} & \textbf{Fine-tuning} & \textbf{Layer Freezing} & \textbf{Without \gls{tl}} \\
    \midrule
    No Lapse & N/A & $0.4894$ & $0.3361$ & N/A \\
    Static Lapse & N/A & $0.0794$ & $0.0763$ & N/A \\
    Dynamic Lapse & $0.0587$ &  N/A &  N/A & $0.2950$ \\
    \bottomrule
    \end{tabular}
    \caption{Comparison of different \gls{tl} methods on \gls{gmmb} contracts}
    \label{tab3:transfer_learning_results}
\end{table}

Table~\ref{tab3:transfer_learning_results} summarizes the true \gls{mse}s of the \gls{lstm} metamodels trained using different \gls{tl} methods across various source tasks.
The calculations are based on the metamodel predictions and the true contract losses approximated with $\num{100000}$ inner replications.
The first two rows show the performance of transferring knowledge from \gls{gmmb} contracts with no lapse and the static lapse to the target task of \gls{gmmb} with the dynamic lapse.
The last row shows the performance of training without \gls{tl} on the target task.
The results demonstrate the effectiveness of fine-tuning and layer freezing in transferring knowledge from a related source tasks to the target task. 
When transferring from a source task with the static lapse to a target task with the dynamic lapse, the MSEs achieved are $0.0794$ for fine-tuning and $0.0763$ for layer freezing.
The benchmark MSE of $0.0587$, obtained from extensive training on the \gls{gmmb} dynamic lapse task with much more samples.

@@

However, when the source task is less similar to the target task, the benefits of \gls{tl} are less pronounced. The MSEs in this case are higher, with fine-tuning resulting in $0.4894$ and layer freezing achieving $0.3361$. 
This suggests that the divergence between the source and target tasks can lead to negative transfer.
These results highlight the importance of selecting source tasks that share significant similarities with the target task to maximize the effectiveness of \gls{tl}. 
When the source and target contracts are substantially different, the pre-trained metamodels may struggle to adapt to the new conditions, as they may not have learned features relevant to the target task.
There is no negative transfer observed, but the knowledge transfer is limited.

\subsection{Transfer Knowledge to other Contract Types}

@@

\begin{figure}[ht!]
    \centering
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{./project3/figures/figure4a.png}
        \caption{Extensive Training on Target Task} 
        \label{subfig3-4:extensive}
    \end{subfigure}\hfill
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{./project3/figures/figure4b.png}
        \caption{Without \gls{tl}}
        \label{subfig3-4:without}
    \end{subfigure}
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{./project3/figures/figure4c.png}
        \caption{With Fine-tuning}
        \label{subfig3-4:fineTuning}
    \end{subfigure}\hfill
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{./project3/figures/figure4d.png}
        \caption{With Layer Freezing}
        \label{subfig3-4:layerFreezing}
    \end{subfigure}
    \caption{TL performance on RS-GBM GMWB with dynamic lapse}
    \label{fig3:figure4}
\end{figure}

Figure~\ref{subfig3-4:extensive} presents the performance of an extensively trained \gls{lstm} metamodel on the target task using $\num{50000}$ samples, serving as a benchmark of the best metamodel performance. 
The extensively trained metamodel achieves a low validation error and demonstrates stable convergence due to having enough training samples. 
In contrast, Figure~\ref{subfig3-4:without} shows the results of training the metamodel directly on the target task with $\num{2000}$ training samples. 
Fewer training samples and no prior knowledge leads to unstable training dynamics. 
The validation MSE fluctuates substantially.

@@

\begin{table}[ht!]
    \centering
    \begin{tabular}{lrrr}
        \toprule
        \textbf{Model} & \textbf{Training samples} & \textbf{Training MSE} & \textbf{True MSE} \\
        \midrule
        Without \gls{tl} & $\num{2000}$ & $0.3588$ & $0.4188$ \\
        Fine-tuning & $\num{2000}$ & $0.1690$ & $0.1780$ \\
        Layer freezing & $\num{2000}$ & $0.1828$ & $0.2295$ \\
        Extensive training & $\num{50000}$ & $0.0853$ & $0.0726$ \\
        \bottomrule
    \end{tabular}
    \caption{Comparison of different TL methods to GMWB contracts}
    \label{tab3:transfer_learning_results_gmwb}
\end{table}

Table~\ref{tab3:transfer_learning_results_gmwb} summarizes the true MSEs of the \gls{lstm} metamodels on the \gls{gmwb} contracts.
Training MSE and true MSE are computed using the training labels and the true contract losses approximated with $\num{100000}$ inner replications, respectively.
This is consistent with the terminology used in Chapter~\ref{chap:project2}.
The \gls{tl} methods are compared to training from the scratch.
The suboptimal performance of layer freezing in this context indicates that the difference between the source and target tasks is too substantial for this method to be effective. 
While layer freezing can be advantageous when the source and target tasks are closely related, it may hinder performance when the tasks diverge substantially.
In such circumstances, fine-tuning provides a better approach by allowing the metamodel to leverage transferable knowledge while adapting to the new task's specific requirements. 
Fine-tuning enables both the \gls{lstm} and the fully connected layers to update their weights, capturing the complex dynamics of the \gls{gmwb} contracts more effectively.
This is particularly beneficial when developing metamodels for the complex \gls{va} contracts with limited simulation data. 
For instance, transferring information from the \gls{gmmb} contracts can still be valuable when modeling the \gls{gmwb} contracts.
Both contracts share some common contract features and temporal patterns, which allows a pre-trained \gls{lstm} metamodel to capture generalizable features that provide a solid foundation for the target task.

\begin{figure}[ht!]
    \centering
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{./project3/figures/figure4_1.png}
        \caption{Identifying Tail scenarios} 
        \label{subfig3-4-1:tail}
    \end{subfigure}\hfill
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{./project3/figures/figure4_2.png}
        \caption{$95\%$-CVaR predictions}
        \label{subfig3-4-2:CVaR}
    \end{subfigure}
    \caption{TL performance on \gls{rsgbm} GMWB with dynamic lapse}
    \label{fig3:figure4-1}
\end{figure}

@@

This finding is consistent with the training history and MSEs in Figure~\ref{fig3:figure4} and Table~\ref{tab3:transfer_learning_results_gmwb}, respectively.
It is important to note that while these results are promising, they also highlight the complexity of modeling the \gls{gmwb} contracts with the dynamic lapse. 
The fact that fine-tuning outperforms layer freezing suggests that there are substantial differences in the tail behavior of the \gls{gmmb} and \gls{gmwb} contracts, particularly when the dynamic lapse is considered. 
These findings highlight the need for careful model selection and validation when applying the \gls{tl} techniques to different \gls{va} products.
When the source and target tasks are substantially different, all layers need to be updated.
Freezing some layers may lead to negative transfer.

@@

Multi-task learning enables the \gls{lstm} metamodels to learn shared representations across related \gls{va} contracts.
In this section, we examine the performance of multi-task learning applied to two types of \gls{va} contracts, \gls{gmmb} and \gls{gmwb} with the dynamic lapse rates.
The simulation datasets contain $2,\!000$ samples for each contract type, and the \gls{lstm} metamodels are trained using multi-task learning. 
In our experiments, Algorithm~\ref{alg3:multiTaskLearning} is used to train the \gls{lstm} metamodels simultaneously to minimize the multi-task MSE loss function in Equation~\eqref{eq3:multiTaskLoss}.
We use individual task training as a baseline for comparison, where the \gls{lstm} metamodels are trained separately on the \gls{gmmb} and \gls{gmwb} contracts.
The objective is to assess how multi-task learning can improve the training efficiency and performance of both products compared to individual task training.

@@

\begin{figure}[ht!]
    \centering
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{./project3/figures/figure5a.png}
        \caption{Multi-task training history} 
        \label{subfig3-5:multiTask}
    \end{subfigure}\hfill
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{./project3/figures/figure5b.png}
        \caption{Task performance with multi-task training}
        \label{subfig3-5:fineTuning}
    \end{subfigure}
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{./project3/figures/figure5c.png}
        \caption{GMMB individual task training} 
    \label{subfig3-5:gmmb_individual}
    \end{subfigure}\hfill
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{./project3/figures/figure5d.png}
        \caption{GMWB individual task training}
        \label{subfig3-5:gmwb_individual}
    \end{subfigure}
    \caption{Multi-task Learning on \gls{rsgbm} GMMB and GMWB with dynamic lapse}
    \label{fig3:figure5}
\end{figure}

>>>>

\begin{algorithm}
    \caption{Multi-task Learning Framework for LSTM Metamodels}
    \begin{algorithmic}[1] \label{alg3:multiTaskLearning}
        \STATE \textbf{Input:} learning rate $\alpha$, set of $K$ tasks $\{\mathcal{T}_k\}_{k=1}^K$ with datasets $\mathcal{D}_k = \{(X_k^{(i)}, L_k^{(i)})\}_{i=1}^{M_k}$, task-specific parameters $\theta_k$ for each task $k$, and shared parameters $\theta_0$.

@@

\begin{algorithm}
    \caption{TL Framework for LSTM Metamodels: Combining Fine-tuning, Layer Freezing, and Multi-task Learning}
    \begin{algorithmic}[1] \label{alg3:combined}
        \STATE \textbf{Input:} Set of $K$ tasks $\{\mathcal{T}_k\}_{k=1}^K$, with datasets $\mathcal{D}_k = \{(X_k^{(i)}, L_k^{(i)})\}_{i=1}^{M_k}$ for each task $k$, target dataset $\mathcal{D}_{\text{Ta}}$, learning rate $\alpha_{\text{So}}$ and $\alpha_{\text{Ta}}$, shared parameters $\theta_0$, task-specific parameters $\theta_k$ for each task $k$.

@@

\begin{table}[ht!] 
    \centering
    \begin{tabular}{lcccc} 
    \toprule
    \textbf{Contract} & \textbf{Asset model} & \textbf{Lapse} & \textbf{$M_{\text{So}}$}  & \textbf{$M_{\text{Ta}}$}\\
    \midrule
    \gls{gmmb} & \gls{rsgbm} & No lapse & $\num{50000}$ & N/A \\
    \gls{gmmb} & \gls{rsgbm} & Static lapse & $\num{50000}$ & $\num{2000}$ \\
    \gls{gmmb} & \gls{rsgbm} & Dynamic lapse & $\num{50000}$ & $\num{2000}$ \\
    \gls{gmwb} & \gls{rsgbm} & Dynamic lapse & N/A & $\num{2000}$ \\
    \bottomrule
    \end{tabular}
    \caption{VA contracts for TL experiments}
    \label{tab3:contracts}
\end{table}

@@

All \gls{lstm} metamodels are evaluated based on their training history graphs for the target task, which plot training and validation \gls{mse} against the number of training epochs.
The training history graphs visualize the learning curves of the \gls{lstm} metamodels, which provide insights into stability and convergence behavior the metamodels.
We measure the generalization performance of the \gls{lstm} metamodels using the true \gls{mse}, which quantifies the accuracy of the \gls{lstm} metamodels in approximating the true inner simulation model of \gls{va} contracts.
The true \gls{mse} is computed by comparing the metamodel predictions with the true contract losses, which are approximated by the standard nested simulation procedure with $\num{100000}$ inner replications.
Hyperparameters and network architectures for the \gls{lstm} metamodels are kept the same as the \gls{lstm} metamodel in Section~\ref{sec2:numerical}. 
Due to a limited amount of computational resources, macro replications are not feasible for the \gls{tl} experiments.

@@

\begin{figure}[ht!]
    \centering
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{./project3/figures/figure1a.png}
        \caption{Extensive training on target task} 
        \label{subfig3-1:extensive}
    \end{subfigure}\hfill
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{./project3/figures/figure1b.png}
        \caption{Without TL}
        \label{subfig3-1:without}
    \end{subfigure}
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{./project3/figures/figure1c.png}
        \caption{With fine-tuning}
        \label{subfig3-1:fineTuning}
    \end{subfigure}\hfill
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{./project3/figures/figure1d.png}
        \caption{With layer freezing}
        \label{subfig3-1:layerFreezing}
    \end{subfigure}
    \caption{Metamodel performance on GMMB with RS-GBM and static lapse}
    \label{fig3:figure1}
\end{figure}

@@

\begin{figure}[ht!]
    \centering
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{./project3/figures/figure2a.png}
        \caption{Fine-tuning from no lapse} 
        \label{subfig3-2:fromNolapse}
    \end{subfigure}\hfill
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{./project3/figures/figure2b.png}
        \caption{Fine-tuning from static lapse}
        \label{subfig3-2:fromLapse}
    \end{subfigure}
    \caption{Fine-tuned metamodel performance on GMMB with RS-GBM and dynamic lapse}
    \label{fig3:figure2}
\end{figure}

@@

\begin{figure}[ht!]
    \centering
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{./project3/figures/figure3a.png}
        \caption{Freezing LSTM Layers} 
        \label{subfig3-3:freezeLSTM}
    \end{subfigure}\hfill
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{./project3/figures/figure3b.png}
        \caption{Freezing the fully connected layer}
        \label{subfig3-3:freezeFC}
    \end{subfigure}
    \caption{Layer freezing on GMMB with RS-GBM and dynamic lapse}
    \label{fig3:figure3}
\end{figure}

@@

\begin{table}[ht!]
    \centering
    \begin{tabular}{lllll}
    \toprule
    \textbf{Lapse type} & \textbf{Extensive} & \textbf{Fine-tuning} & \textbf{Layer freezing} & \textbf{Without \gls{tl}} \\
    \midrule
    No Lapse & N/A & $0.4894$ & $0.3361$ & N/A \\
    Static Lapse & N/A & $0.0794$ & $0.0763$ & N/A \\
    Dynamic Lapse & $0.0587$ &  N/A &  N/A & $0.2950$ \\
    \bottomrule
    \end{tabular}
    \caption{Comparison of different TL methods on GMMB contracts}
    \label{tab3:transfer_learning_results}
\end{table}

Table~\ref{tab3:transfer_learning_results} summarizes the true \gls{mse}s of the \gls{lstm} metamodels trained using different \gls{tl} methods across various source tasks.
The calculations are based on the metamodel predictions and the true contract losses approximated with $\num{100000}$ inner replications.
The first two rows show the performance of transferring knowledge from \gls{gmmb} contracts with no lapse and the static lapse to the target task of \gls{gmmb} with the dynamic lapse.
The last row shows the performance of training without \gls{tl} on the target task.
The results demonstrate the effectiveness of fine-tuning and layer freezing in transferring knowledge from a related source tasks to the target task. 
When transferring from a source task with the static lapse to a target task with the dynamic lapse, the \gls{mse}s achieved are $0.0794$ for fine-tuning and $0.0763$ for layer freezing.
The benchmark \gls{mse} of $0.0587$, obtained from extensive training on the \gls{gmmb} dynamic lapse task with much more samples.

@@

However, when the source task is less similar to the target task, the benefits of \gls{tl} are less pronounced. The \gls{mse}s in this case are higher, with fine-tuning resulting in $0.4894$ and layer freezing achieving $0.3361$. 
This suggests that the divergence between the source and target tasks can lead to negative transfer.
These results highlight the importance of selecting source tasks that share significant similarities with the target task to maximize the effectiveness of \gls{tl}. 
When the source and target contracts are substantially different, the pre-trained metamodels may struggle to adapt to the new conditions, as they may not have learned features relevant to the target task.
There is no negative transfer observed, but the knowledge transfer is limited.

\subsection{Transfer Knowledge to Other Contract Types}

@@

\begin{figure}[ht!]
    \centering
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{./project3/figures/figure4a.png}
        \caption{Extensive training on target task} 
        \label{subfig3-4:extensive}
    \end{subfigure}\hfill
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{./project3/figures/figure4b.png}
        \caption{Without TL}
        \label{subfig3-4:without}
    \end{subfigure}
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{./project3/figures/figure4c.png}
        \caption{With fine-tuning}
        \label{subfig3-4:fineTuning}
    \end{subfigure}\hfill
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{./project3/figures/figure4d.png}
        \caption{With layer freezing}
        \label{subfig3-4:layerFreezing}
    \end{subfigure}
    \caption{TL performance on GMWB with RS-GBM and dynamic lapse}
    \label{fig3:figure4}
\end{figure}

Figure~\ref{subfig3-4:extensive} presents the performance of an extensively trained \gls{lstm} metamodel on the target task using $\num{50000}$ samples, serving as a benchmark of the best metamodel performance. 
The extensively trained metamodel achieves a low validation error and demonstrates stable convergence due to having enough training samples. 
In contrast, Figure~\ref{subfig3-4:without} shows the results of training the metamodel directly on the target task with $\num{2000}$ training samples. 
Fewer training samples and no prior knowledge leads to unstable training dynamics. 
The validation \gls{mse} fluctuates substantially.

@@

\begin{table}[ht!]
    \centering
    \begin{tabular}{lrrr}
        \toprule
        \textbf{Model} & \textbf{Training samples} & \textbf{Training \gls{mse}} & \textbf{True \gls{mse}} \\
        \midrule
        Without \gls{tl} & $\num{2000}$ & $0.3588$ & $0.4188$ \\
        Fine-tuning & $\num{2000}$ & $0.1690$ & $0.1780$ \\
        Layer freezing & $\num{2000}$ & $0.1828$ & $0.2295$ \\
        Extensive training & $\num{50000}$ & $0.0853$ & $0.0726$ \\
        \bottomrule
    \end{tabular}
    \caption{Comparison of different TL methods to GMWB contracts}
    \label{tab3:transfer_learning_results_gmwb}
\end{table}

Table~\ref{tab3:transfer_learning_results_gmwb} summarizes the true \gls{mse}s of the \gls{lstm} metamodels on the \gls{gmwb} contracts.
Training \gls{mse} and true \gls{mse} are computed using the training labels and the true contract losses approximated with $\num{100000}$ inner replications, respectively.
This is consistent with the terminology used in Chapter~\ref{chap:project2}.
The \gls{tl} methods are compared to training from the scratch.
The suboptimal performance of layer freezing in this context indicates that the difference between the source and target tasks is too substantial for this method to be effective. 
While layer freezing can be advantageous when the source and target tasks are closely related, it may hinder performance when the tasks diverge substantially.
In such circumstances, fine-tuning provides a better approach by allowing the metamodel to leverage transferable knowledge while adapting to the new task's specific requirements. 
Fine-tuning enables both the \gls{lstm} and the fully connected layers to update their weights, capturing the complex dynamics of the \gls{gmwb} contracts more effectively.
This is particularly beneficial when developing metamodels for the complex \gls{va} contracts with limited simulation data. 
For instance, transferring information from the \gls{gmmb} contracts can still be valuable when modeling the \gls{gmwb} contracts.
Both contracts share some common contract features and temporal patterns, which allows a pre-trained \gls{lstm} metamodel to capture generalizable features that provide a solid foundation for the target task.

\begin{figure}[ht!]
    \centering
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{./project3/figures/figure4_1.png}
        \caption{Identifying tail scenarios} 
        \label{subfig3-4-1:tail}
    \end{subfigure}\hfill
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{./project3/figures/figure4_2.png}
        \caption{$95\%$-CVaR predictions}
        \label{subfig3-4-2:CVaR}
    \end{subfigure}
    \caption{TL performance on GMWB with RS-GBM and dynamic lapse}
    \label{fig3:figure4-1}
\end{figure}

@@

This finding is consistent with the training history and \gls{mse}s in Figure~\ref{fig3:figure4} and Table~\ref{tab3:transfer_learning_results_gmwb}, respectively.
It is important to note that while these results are promising, they also highlight the complexity of modeling the \gls{gmwb} contracts with the dynamic lapse. 
The fact that fine-tuning outperforms layer freezing suggests that there are substantial differences in the tail behavior of the \gls{gmmb} and \gls{gmwb} contracts, particularly when the dynamic lapse is considered. 
These findings highlight the need for careful model selection and validation when applying the \gls{tl} techniques to different \gls{va} products.
When the source and target tasks are substantially different, all layers need to be updated.
Freezing some layers may lead to negative transfer.

@@

Multi-task learning enables the \gls{lstm} metamodels to learn shared representations across related \gls{va} contracts.
In this section, we examine the performance of multi-task learning applied to two types of \gls{va} contracts, \gls{gmmb} and \gls{gmwb} with the dynamic lapse rates.
The simulation datasets contain $2,\!000$ samples for each contract type, and the \gls{lstm} metamodels are trained using multi-task learning. 
In our experiments, Algorithm~\ref{alg3:multiTaskLearning} is used to train the \gls{lstm} metamodels simultaneously to minimize the multi-task \gls{mse} loss function in Equation~\eqref{eq3:multiTaskLoss}.
We use individual task training as a baseline for comparison, where the \gls{lstm} metamodels are trained separately on the \gls{gmmb} and \gls{gmwb} contracts.
The objective is to assess how multi-task learning can improve the training efficiency and performance of both products compared to individual task training.

@@

\begin{figure}[ht!]
    \centering
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{./project3/figures/figure5a.png}
        \caption{Multi-task training history} 
        \label{subfig3-5:multiTask}
    \end{subfigure}\hfill
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{./project3/figures/figure5b.png}
        \caption{Task performance with multi-task training}
        \label{subfig3-5:fineTuning}
    \end{subfigure}
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{./project3/figures/figure5c.png}
        \caption{GMMB individual task training} 
    \label{subfig3-5:gmmb_individual}
    \end{subfigure}\hfill
    \begin{subfigure}{0.48\textwidth}
        \includegraphics[width=\textwidth]{./project3/figures/figure5d.png}
        \caption{GMWB individual task training}
        \label{subfig3-5:gmwb_individual}
    \end{subfigure}
    \caption{Multi-task learning on GMMB and GMWB with RS-GBM and dynamic lapse}
    \label{fig3:figure5}
\end{figure}
